---
title: "Week 6"
author: "Evan Wu"
output: 
  html_document:
    code_folding:
      hide
---

<br>

#### Code Background

Looking at whiff data for right-handed pitchers (from 2022, 85 pitchers with
1842+ seasonal pitches). 

<br>

#### Libraries
```{r Libraries, message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(glmnet)
library(patchwork)
library(olsrr)
library(broom)
```


``` {r Functions, }
# Geom Zone (Jackie's) ####

geom_zone <- function(top = 11/3, bottom = 3/2, linecolor = "black"){
  geom_rect(xmin = -.7083, xmax = .7083, ymin = bottom, ymax = top,
            alpha = 0, color = linecolor, linewidth = 0.75)
}

# c(0, 0, -.25, -.5, -.25))

geom_plate <- function(pov = "pitcher"){
   df <- case_when(
     pov == "pitcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, .25, .5, .25)),
     pov == "catcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, -.25, -.5, -.25))
   )
  
   g <- geom_polygon(data = df, aes(x = x, y = y), fill = "white", color = "black", linewidth = 1.25)
  g
}

# Barrel Function
is.barrel <- function(LA, EV){
  upper <- 1.11*EV - 78.89
  lower <- -EV + 124
  outcome <- (LA >= lower) & (LA <= upper) & (EV >= 98) & (LA >= 8) & (LA <= 50)
  outcome <- replace_na(outcome, FALSE)
  outcome
}

```

#### Data
Added code to calculate whether a batted ball is a barrel or not.
Used code to find the slopes and intercepts for both lines and then inputted
rounded coefficients into the model for more simplicity.

Added predicted whiff rate. There were a few hitters with more than 300 pitches
of a given pitch type, so for those I used their true whiff percentage.

Also added a strike metric that includes all pitches that can strike a hitter out:
called and swinging strikes, as well as foul tips and foul bunts.

```{r Data, message = FALSE, warning = FALSE}

rhp <- read_csv("CSVs/rhp_pitches.csv") %>% 
  select(-...1) %>% 
  filter(!is.na(pfx_z)) %>% 
  mutate(distance_sweet = sqrt(((plate_x - 0.85)^2)+((plate_z - 1.55)^2))) %>% 
  mutate(zone = case_when(distance < 2 ~ zone,
                          distance >= 2 & plate_z > (sz_top + sz_bot)/2 ~ 16,
                          distance >= 2 & plate_z <= (sz_top + sz_bot)/2 ~ 17)) %>% 
  mutate(ab_id = paste0(game_date, "_", player_id, "_", at_bat_number),
         prev_ab_id = lead(ab_id, 1)) %>% 
  mutate(prev_pitch = ifelse(ab_id == prev_ab_id, lead(pitch_type, 1), NA)) %>% 
  select(-ab_id, -prev_ab_id) %>% 
  mutate(prev_pitch = case_when(prev_pitch  %in% c("FF", "FC", "SI") ~ "Fastball",
                                prev_pitch %in% c("SL", "CU", "KC", 
                                                  "SV", "ST", "CS") ~ "Breaking Ball",
                                prev_pitch  %in% c("CH", "FS") ~ "Off Speed")) %>% 
  mutate(zone = relevel(as.factor(zone), ref = 5))

# New Code for Barrels

# Upper Limit
m <- (50 - 30) / (116 - 98)
b <- 30 - m * 98
paste0("y = ", round(m, 2), "x ", round(b, 2))

# Lower Limit
m <- (8 - 26) / (116 - 98)
b <- 26 - m * 98
paste0("y = ", round(m, 2), "x + ", round(b, 2))

# Barrel
rhp <- rhp %>% 
  mutate(Barrel = is.barrel(LA = launch_angle, EV = launch_speed))

# Strike
rhp <- rhp %>% 
  mutate(STRIKE = ifelse(description  %in% c("called_strike", "swinging_strike",
                                           "foul_tip", "bunt_foul_tip", "foul_bunt",
                                           "swinging_strike_blocked", "missed_bunt"),
                       1, 0))


# Batter Whiff
batter_stats <- rhp %>% 
  summarize(rate = mean(whiff),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_whiff = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bwhiff = (pitches / 300)*rate + ((300-pitches)/300)*pitch_whiff) %>% 
  mutate(pred_bwhiff = ifelse(pitches >= 300, rate, pred_bwhiff))


# Merging Whiff Prediction with RHP
rhp <- rhp %>% 
  left_join(select(batter_stats, batter, pitch_type, pred_bwhiff), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))

rhp <- rhp %>% 
  mutate(prev_pitch = ifelse(is.na(prev_pitch), "None", prev_pitch)) %>% 
  mutate(Count = paste0(balls, "-", strikes)) %>% 
  mutate(Count = as.factor(Count))


# Barrel
batter_stats2 <- rhp %>% 
  summarize(rate = mean(Barrel),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_barrel = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bbarrel = (pitches / 300)*rate + ((300-pitches)/300)*pitch_barrel) %>% 
  mutate(pred_bbarrel = ifelse(pitches >= 300, rate, pred_bbarrel))

rhp <- rhp %>% 
  left_join(select(batter_stats2, batter, pitch_type, pred_bbarrel), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))


sliders <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == "SL") %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch)) %>% 
  mutate(prev_pitch_ff = ifelse(prev_pitch == "FF", 1, 0))
  
fastballs <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == fb_type) %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch))

```



#### WHIFF Slider Refined Model & Predictions

Variables Removed:

count

plate_x^2

pfx_x_fb_diff

pfx_z_fb_diff

distance


```{r Slider Refined Model Whiff, message = FALSE}
# Refined Model
sl_model_whiff <- glm(whiff ~ pitch_speed + pfx_x + pfx_z + plate_x + plate_z + zone +
              pfx_total + speed_change + break_change + prev_pitch +
                speed_fb_diff + release_spin_rate +
                I(plate_z^2) + pred_bwhiff,
              data = sliders, family = binomial)

# Model Evaluation
summary(sl_model_whiff)
sl_whiff <- augment(sl_model_whiff)
  
sl_whiff <- sl_whiff %>% 
  mutate(prediction_log = predict(sl_model_whiff, sl_whiff),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "slider")

```

Slider Refined Model: 

The model seems to be accurate the expected proportion of the time, except after
around 0.4. Coincidentally, this is around where the observations thin out.
This could possibly be a result of a lack of datapoints?





#### WHIFF Fastball Refined Model & Predictions

Variables Removed:

count

plate_x

distance

plate_x +

```{r Fastball Refined Model Whiff, message = FALSE}
# Refined Model
fb_model_whiff <- glm(whiff ~ pitch_speed + pfx_x + pfx_z + plate_x + plate_z + zone +
              pfx_total + speed_change + break_change + prev_pitch +
                release_spin_rate + I(plate_x^2) + I(plate_z^2) + pred_bwhiff,
              data = fastballs, family = binomial)

# Model Evaluation
summary(fb_model_whiff)

fb_whiff <- augment(fb_model_whiff)
  
fb_whiff <- fb_whiff %>% 
  mutate(prediction_log = predict(fb_model_whiff, fb_whiff),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "fastball")

```
Refined Fastball Model:

It is interesting that every variable besides the count variable was significant.
The model also diverges from the expected prediction accuracy when the amount
of observations thins.



#### BARREL Slider Refined Model & Predictions

Variables Removed:

count

pfx_x_fb_diff

release_spin_rate

pfx_z_fb_diff

pfx_total

pfx_x

pfx_z

speed_change

plate_x

break_change

plate_x +


```{r Slider Refined Model Barrel, message = FALSE}
# Refined Model
sl_model_barrel <- glm(Barrel ~ pitch_speed + plate_z + zone + plate_x +
              prev_pitch + speed_fb_diff + I(plate_x^2) + I(plate_z^2) + pred_bbarrel,
              data = sliders, family = binomial)


# Model Evaluation
summary(sl_model_barrel)
sl_barrel <- augment(sl_model_barrel)
  
sl_barrel <- sl_barrel %>% 
  mutate(prediction_log = predict(sl_model_barrel, sl_barrel),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "barrel",
         pitch = "slider")

```
Refined Slider Model: 

The prediction seems to underpredict barrel rate. This might be because
barrels are so rare with each observation as a pitch.




#### BARREL Fastball Refined Model & Predictions

Variables Removed:

count

pfx_total

plate_x

pfx_x

distance

plate_x +



```{r Fastball Refined Model Barrel, message = FALSE}
# Refined Model
fb_model_barrel <- glm(Barrel ~ pitch_speed + pfx_z + plate_z + zone +
              break_change + prev_pitch + plate_x +
                release_spin_rate + I(plate_x^2) + I(plate_z^2) + pred_bbarrel,
              data = fastballs, family = binomial)


# Model Evaluation
summary(fb_model_barrel)
fb_barrel <- augment(fb_model_barrel)
  
fb_barrel <- fb_barrel %>% 
  mutate(prediction_log = predict(fb_model_barrel, fb_barrel),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "barrel",
         pitch = "fastball")
```
Refined Fastball Model:

For both barrel models, adding predicted barrel whiff (of batter) replaces
distance. Not exactly sure why. The fastball model also had speed_change
supplanted by the barrel whiff variable.
This model seems get underpredict by slightly more as barrel rate/predicted
barrel rate increases (linearly), while the smooth function severly depreciates in
accuracy around 0.15 barrel rate. 



#### STRIKE Slider Refined Model & Predictions

Variables Removed:

pfx_x_fb_diff

pitch_speed

pfx_z_fb_diff

speed_change


```{r Slider Refined Model Strike, message = FALSE}
# Refined Model

sliders <- sliders %>% 
  mutate(fb_prev = ifelse(prev_pitch == "Fastball", 1, 0)) %>% 
  mutate(count_s = case_when(count  %in% c("0_0", "1_1", "2_2") ~ "Even",
                             count  %in%  c("1_0", "2_1", "2_0", "3_0", "3_1") ~ "Hitter",
                             count %in%  c("0_1", "0_2", "1_2") ~ "Pitcher",
                             count == "3_2" ~ "Full"))

sl_model_strike <- glm(STRIKE ~ pfx_x + pfx_z + plate_x + plate_z + zone +
              count_s + pfx_total + break_change + distance + prev_pitch +
                speed_fb_diff + release_spin_rate +
                I(plate_x^2) + I(plate_z^2),
              data = sliders, family = binomial)

# Model Evaluation
summary(sl_model_strike)
sl_strike <- augment(sl_model_strike)
  
sl_strike <- sl_strike %>% 
  mutate(prediction_log = predict(sl_model_strike, sl_strike),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "strike",
         pitch = "slider")
```
Slider Refined Model:

This model also roughly predicts strike rate. As the data thins out,
the model seems to slightly overpredict strikes. There is also a weird
dip in predictions from around 0.3 to 0.4, which is intriguing.



#### STRIKE Fastball Refined Model & Predictions

Variables Removed:

count

prev_pitch

speed_change

```{r Fastball Refined Model Strike, message = FALSE}
# Refined Model

fb_model_strike <- glm(STRIKE ~ pitch_speed + pfx_x + pfx_z + plate_x + plate_z + zone +
              count + pfx_total + break_change + distance + 
                release_spin_rate + I(plate_x^2) + I(plate_z^2),
              data = fastballs, family = binomial)


# Model Evaluation
summary(fb_model_strike)
fb_strike <- augment(fb_model_strike)
  
fb_strike <- fb_strike %>% 
  mutate(prediction_log = predict(fb_model_strike, fb_strike),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "strike",
         pitch = "fastball")

```

Fastball Refined Model:

This model seems to be pretty accurate, including extrapolation. There is a similar,
albeit slightly less pronounced, dip in prediction values between 0.3 and 0.4.


#### Making Naive Models

```{r Naive Models}

## Intercept-Only Models

# Slider

sl_model_whiff_int <- glm(whiff ~ 1, 
                      data = sliders, family = binomial) 
sl_whiff_int <- augment(sl_model_whiff_int)


sl_whiff_int <- sl_whiff_int %>% 
  mutate(prediction_log = predict(sl_model_whiff_int, sl_whiff_int),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "slider",
         type = "intercept only")


# Fastball

fb_model_whiff_int <- glm(whiff ~ 1, 
                      data = fastballs, family = binomial) 
fb_whiff_int <- augment(fb_model_whiff_int)


fb_whiff_int <- fb_whiff_int %>% 
  mutate(prediction_log = predict(fb_model_whiff_int, fb_whiff_int),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "fastball",
         type = "intercept only")


## Zone-Only Models

# Slider

sl_model_whiff_zone <- glm(whiff ~ zone, 
                      data = sliders, family = binomial) 

sl_whiff_zone <- augment(sl_model_whiff_zone)


sl_whiff_zone <- sl_whiff_zone %>% 
  mutate(prediction_log = predict(sl_model_whiff_zone, sl_whiff_zone),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "slider",
         type = "zone only")


# Fastball

fb_model_whiff_zone <- glm(whiff ~ zone, 
                      data = fastballs, family = binomial) 

fb_whiff_zone <- augment(fb_model_whiff_zone)


fb_whiff_zone <- fb_whiff_zone %>% 
  mutate(prediction_log = predict(fb_model_whiff_zone, fb_whiff_zone),
         prediction = 1 / (1 + exp(-prediction_log)),
         rounded_pred = case_when(
           prediction >= 0.5 ~ 1,
           prediction < 0.5 ~ 0
         )) %>% 
  filter(!is.na(prediction)) %>% 
  mutate(Response = "whiff",
         pitch = "fastball",
         type = "zone only")


```


#### Combining All Models

```{r Combining All Augmented/Predicted}

all_models_data <- bind_rows(
                    sl_whiff, fb_whiff,
                    sl_barrel, fb_barrel,
                    sl_strike, fb_strike,
                    sl_whiff_int, fb_whiff_int,
                    sl_whiff_zone, fb_whiff_zone) %>% 
  select(Response:pitch, type, whiff:pred_bwhiff, `I(plate_x^2)`:count, `.fitted`:rounded_pred)

all_models <- bind_rows(
  tidy(sl_model_whiff) %>% mutate(Response = "whiff", pitch = "slider", type = "complex"), 
  tidy(fb_model_whiff) %>% mutate(Response = "whiff", pitch = "fastball", type = "complex"),
  tidy(sl_model_barrel) %>% mutate(Response = "barrel", pitch = "slider", type = "complex"),
  tidy(fb_model_barrel) %>% mutate(Response = "barrel", pitch = "fastball", type = "complex"),
  tidy(sl_model_strike) %>% mutate(Response = "strike", pitch = "slider", type = "complex"),
  tidy(fb_model_strike) %>% mutate(Response = "strike", pitch = "fastball", type = "complex"),
  tidy(sl_model_whiff_int) %>% mutate(Response = "whiff", pitch = "slider", 
                                      type = "intercept only"),
  tidy(fb_model_whiff_int) %>% mutate(Response = "whiff", pitch = "fastball", 
                                      type = "intercept only"),
  tidy(sl_model_whiff_zone) %>% mutate(Response = "whiff", pitch = "slider", 
                                      type = "zone only"),
  tidy(fb_model_whiff_zone) %>% mutate(Response = "whiff", pitch = "fastball", 
                                      type = "zone only")
)

```

