---
title: "Week 8"
author: "Evan Wu"
output: 
  html_document:
    code_folding:
      hide
---

<br>

#### Code Background

Looking at whiff data for right-handed pitchers (from 2022, 85 pitchers with
1842+ seasonal pitches). 

<br>

#### Libraries
```{r Libraries, message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(glmnet)
library(patchwork)
library(olsrr)
library(broom)
options(scipen = 999)
library(RColorBrewer)
```


``` {r Functions, }
# Geom Zone (Jackie's) ####

geom_zone <- function(top = 11/3, bottom = 3/2, linecolor = "black"){
  geom_rect(xmin = -.7083, xmax = .7083, ymin = bottom, ymax = top,
            alpha = 0, color = linecolor, linewidth = 0.75)
}

# c(0, 0, -.25, -.5, -.25))

geom_plate <- function(pov = "pitcher"){
   df <- case_when(
     pov == "pitcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, .25, .5, .25)),
     pov == "catcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, -.25, -.5, -.25))
   )
  
   g <- geom_polygon(data = df, aes(x = x, y = y), fill = "white", color = "black", linewidth = 1.25)
  g
}

# Barrel Function
is.barrel <- function(LA, EV){
  upper <- 1.11*EV - 78.89
  lower <- -EV + 124
  outcome <- (LA >= lower) & (LA <= upper) & (EV >= 98) & (LA >= 8) & (LA <= 50)
  outcome <- replace_na(outcome, FALSE)
  outcome
}

```

#### Data
Added code to calculate whether a batted ball is a barrel or not.
Used code to find the slopes and intercepts for both lines and then inputted
rounded coefficients into the model for more simplicity.

Added predicted whiff rate. There were a few hitters with more than 300 pitches
of a given pitch type, so for those I used their true whiff percentage.

Also added a strike metric that includes all pitches that can strike a hitter out:
called and swinging strikes, as well as foul tips and foul bunts.

```{r Data, message = FALSE, warning = FALSE}

rhp <- read_csv("CSVs/rhp_pitches.csv") %>% 
  select(-...1) %>% 
  filter(!is.na(pfx_z)) %>% 
  mutate(distance_sweet = sqrt(((plate_x - 0.85)^2)+((plate_z - 1.55)^2))) %>% 
  mutate(zone = case_when(distance < 2 ~ zone,
                          distance >= 2 & plate_z > (sz_top + sz_bot)/2 ~ 16,
                          distance >= 2 & plate_z <= (sz_top + sz_bot)/2 ~ 17)) %>% 
  mutate(ab_id = paste0(game_date, "_", player_id, "_", at_bat_number),
         prev_ab_id = lead(ab_id, 1)) %>% 
  mutate(prev_pitch = ifelse(ab_id == prev_ab_id, lead(pitch_type, 1), NA)) %>% 
  select(-ab_id, -prev_ab_id) %>% 
  mutate(prev_pitch = case_when(prev_pitch  %in% c("FF", "FC", "SI") ~ "Fastball",
                                prev_pitch %in% c("SL", "CU", "KC", 
                                                  "SV", "ST", "CS") ~ "Breaking Ball",
                                prev_pitch  %in% c("CH", "FS") ~ "Off Speed")) %>% 
  mutate(zone = relevel(as.factor(zone), ref = 5))

# New Code for Barrels

# Upper Limit
m <- (50 - 30) / (116 - 98)
b <- 30 - m * 98
paste0("y = ", round(m, 2), "x ", round(b, 2))

# Lower Limit
m <- (8 - 26) / (116 - 98)
b <- 26 - m * 98
paste0("y = ", round(m, 2), "x + ", round(b, 2))

# Barrel
rhp <- rhp %>% 
  mutate(Barrel = is.barrel(LA = launch_angle, EV = launch_speed))

# Strike
rhp <- rhp %>% 
  mutate(STRIKE = ifelse(description  %in% c("called_strike", "swinging_strike",
                                           "foul_tip", "bunt_foul_tip", "foul_bunt",
                                           "swinging_strike_blocked", "missed_bunt"),
                       1, 0))


# Batter Whiff
batter_stats <- rhp %>% 
  summarize(rate = mean(whiff),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_whiff = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bwhiff = (pitches / 300)*rate + ((300-pitches)/300)*pitch_whiff) %>% 
  mutate(pred_bwhiff = ifelse(pitches >= 300, rate, pred_bwhiff))


# Merging Whiff Prediction with RHP
rhp <- rhp %>% 
  left_join(select(batter_stats, batter, pitch_type, pred_bwhiff), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))

rhp <- rhp %>% 
  mutate(prev_pitch = ifelse(is.na(prev_pitch), "None", prev_pitch)) %>% 
  mutate(Count = paste0(balls, "-", strikes)) %>% 
  mutate(Count = as.factor(Count))

# Adding Proportional Distance
rhp <- rhp %>% 
  mutate(dist_x = plate_x/0.708333,
         center = (sz_top + sz_bot) / 2,
         dist_z = ifelse(plate_z >= center, 
                         (plate_z - center) / (sz_top - center), 
                         (plate_z - center) / (center - sz_bot)),
         dist_prop = sqrt(dist_z^2 + dist_x^2))


# Barrel
batter_stats2 <- rhp %>% 
  summarize(rate = mean(Barrel),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_barrel = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bbarrel = (pitches / 300)*rate + ((300-pitches)/300)*pitch_barrel) %>% 
  mutate(pred_bbarrel = ifelse(pitches >= 300, rate, pred_bbarrel))

rhp <- rhp %>% 
  left_join(select(batter_stats2, batter, pitch_type, pred_bbarrel), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))


sliders <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == "SL") %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch)) %>% 
  mutate(prev_pitch_ff = ifelse(prev_pitch == "FF", 1, 0))
  
fastballs <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == fb_type) %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch))

```


```{r}
rhp <- read_csv("CSVs/rhp_pitches.csv") %>% 
  select(-...1)

lhp <- read_csv("CSVs/lhp_pitches.csv") %>% 
  select(-...1)

data <- rbind(rhp, lhp)

data %>% 
  mutate(pitch_speed = round(pitch_speed, 0)) %>% 
  group_by(p_throws, hitter, pitch_type, pitch_speed) %>%
  filter(!is.na(pfx_x)) %>% 
  filter(!is.na(pfx_z)) %>% 
  summarize(min_x = min(pfx_x), max_x = max(pfx_x),
            min_z = min(pfx_z), max_z = max(pfx_z),
            min_spin = min(release_spin_rate, na.rm = TRUE),
            max_spin = max(release_spin_rate, na.rm = TRUE)) %>% 
  write.csv("Movement.csv")
```




