---
title: "Week 9"
author: "Evan Wu"
output: 
  html_document:
    code_folding:
      hide
---

<br>

#### Code Background

Looking at whiff data for right-handed pitchers (from 2022, 85 pitchers with
1842+ seasonal pitches). 

<br>

#### Libraries
```{r Libraries, message = FALSE, warning = FALSE}
library(tidyverse)
library(knitr)
library(glmnet)
library(patchwork)
library(olsrr)
library(broom)
options(scipen = 999)
library(RColorBrewer)
library(Metrics)
```


``` {r Functions, }
# Geom Zone (Jackie's) ####

geom_zone <- function(top = 11/3, bottom = 3/2, linecolor = "black"){
  geom_rect(xmin = -.7083, xmax = .7083, ymin = bottom, ymax = top,
            alpha = 0, color = linecolor, linewidth = 0.75)
}

# c(0, 0, -.25, -.5, -.25))

geom_plate <- function(pov = "pitcher"){
   df <- case_when(
     pov == "pitcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, .25, .5, .25)),
     pov == "catcher" ~ 
       data.frame(x = c(-.7083, .7083, .7083 ,0, -.7083), y = c(0, 0, -.25, -.5, -.25))
   )
  
   g <- geom_polygon(data = df, aes(x = x, y = y), fill = "white", color = "black", linewidth = 1.25)
  g
}

# Barrel Function
is.barrel <- function(LA, EV){
  upper <- 1.11*EV - 78.89
  lower <- -EV + 124
  outcome <- (LA >= lower) & (LA <= upper) & (EV >= 98) & (LA >= 8) & (LA <= 50)
  outcome <- replace_na(outcome, FALSE)
  outcome
}

swap_names <- function(name) {
  parts <- strsplit(name, ", ")[[1]]
  if (length(parts) == 2) {
    return(paste(rev(parts), collapse = " "))
  } else {
    return(name)
  }
}

```

#### Data
Added code to calculate whether a batted ball is a barrel or not.
Used code to find the slopes and intercepts for both lines and then inputted
rounded coefficients into the model for more simplicity.

Added predicted whiff rate. There were a few hitters with more than 300 pitches
of a given pitch type, so for those I used their true whiff percentage.

Also added a strike metric that includes all pitches that can strike a hitter out:
called and swinging strikes, as well as foul tips and foul bunts.

```{r Data, message = FALSE, warning = FALSE}

rhp <- read_csv("CSVs/rhp_pitches.csv") %>% 
  select(-...1) %>% 
  filter(!is.na(pfx_z)) %>% 
  mutate(distance_sweet = sqrt(((plate_x - 0.85)^2)+((plate_z - 1.55)^2))) %>% 
  filter(distance < 2) %>% 
  mutate(ab_id = paste0(game_date, "_", player_id, "_", at_bat_number),
         prev_ab_id = lead(ab_id, 1)) %>% 
  mutate(prev_pitch = ifelse(ab_id == prev_ab_id, lead(pitch_type, 1), NA)) %>% 
  select(-ab_id, -prev_ab_id) %>% 
  mutate(prev_pitch = case_when(prev_pitch  %in% c("FF", "FC", "SI") ~ "Fastball",
                                prev_pitch %in% c("SL", "CU", "KC", 
                                                  "SV", "ST", "CS") ~ "Breaking Ball",
                                prev_pitch  %in% c("CH", "FS") ~ "Off Speed")) %>% 
  mutate(zone = relevel(as.factor(zone), ref = 5))


# Barrel
rhp <- rhp %>% 
  mutate(Barrel = is.barrel(LA = launch_angle, EV = launch_speed))

# Strike
rhp <- rhp %>% 
  mutate(STRIKE = ifelse(description  %in% c("called_strike", "swinging_strike",
                                           "foul_tip", "bunt_foul_tip", "foul_bunt",
                                           "swinging_strike_blocked", "missed_bunt"),
                       1, 0))


# Batter Whiff
batter_stats <- rhp %>% 
  summarize(rate = mean(whiff),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_whiff = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bwhiff = (pitches / 300)*rate + ((300-pitches)/300)*pitch_whiff) %>% 
  mutate(pred_bwhiff = ifelse(pitches >= 300, rate, pred_bwhiff))


# Merging Whiff Prediction with RHP
rhp <- rhp %>% 
  left_join(select(batter_stats, batter, pitch_type, pred_bwhiff), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))

rhp <- rhp %>% 
  mutate(prev_pitch = ifelse(is.na(prev_pitch), "None", prev_pitch)) %>% 
  mutate(Count = paste0(balls, "-", strikes)) %>% 
  mutate(Count = as.factor(Count))

# Adding Proportional Distance
rhp <- rhp %>% 
  mutate(dist_x = plate_x/0.708333,
         center = (sz_top + sz_bot) / 2,
         dist_z = ifelse(plate_z >= center, 
                         (plate_z - center) / (sz_top - center), 
                         (plate_z - center) / (center - sz_bot)),
         dist_prop = sqrt(dist_z^2 + dist_x^2))


# Barrel
batter_stats2 <- rhp %>% 
  summarize(rate = mean(Barrel),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_barrel = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bbarrel = (pitches / 300)*rate + ((300-pitches)/300)*pitch_barrel) %>% 
  mutate(pred_bbarrel = ifelse(pitches >= 300, rate, pred_bbarrel))

rhp <- rhp %>% 
  left_join(select(batter_stats2, batter, pitch_type, pred_bbarrel), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))


# Strike
batter_stats3 <- rhp %>% 
  summarize(rate = mean(STRIKE),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_strike = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bstrike = (pitches / 300)*rate + ((300-pitches)/300)*pitch_strike) %>% 
  mutate(pred_bstrike = ifelse(pitches >= 300, rate, pred_bstrike))

rhp <- rhp %>% 
  left_join(select(batter_stats3, batter, pitch_type, pred_bstrike), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))


sliders <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == "SL") %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch)) %>% 
  mutate(prev_pitch_ff = ifelse(prev_pitch == "FF", 1, 0))
  
fastballs <- rhp %>% 
  filter(hitter == "R") %>% 
  filter(pitch_type == fb_type) %>% 
  mutate(count = paste0(balls, "_", strikes),
         prev_pitch = as.factor(prev_pitch))



lhp <- read_csv("CSVs/lhp_pitches.csv") %>% 
  select(-...1) %>% 
  filter(!is.na(pfx_z)) %>% 
  mutate(distance_sweet = sqrt(((plate_x - 0.85)^2)+((plate_z - 1.55)^2))) %>% 
  mutate(zone = case_when(distance < 2 ~ zone,
                          distance >= 2 & plate_z > (sz_top + sz_bot)/2 ~ 16,
                          distance >= 2 & plate_z <= (sz_top + sz_bot)/2 ~ 17)) %>% 
  mutate(ab_id = paste0(game_date, "_", player_id, "_", at_bat_number),
         prev_ab_id = lead(ab_id, 1)) %>% 
  mutate(prev_pitch = ifelse(ab_id == prev_ab_id, lead(pitch_type, 1), NA)) %>% 
  select(-ab_id, -prev_ab_id) %>% 
  mutate(prev_pitch = case_when(prev_pitch  %in% c("FF", "FC", "SI") ~ "Fastball",
                                prev_pitch %in% c("SL", "CU", "KC", 
                                                  "SV", "ST", "CS") ~ "Breaking Ball",
                                prev_pitch  %in% c("CH", "FS") ~ "Off Speed")) %>% 
  mutate(zone = relevel(as.factor(zone), ref = 5))

# Barrel
lhp <- lhp %>% 
  mutate(Barrel = is.barrel(LA = launch_angle, EV = launch_speed))

# Strike
lhp <- lhp %>% 
  mutate(STRIKE = ifelse(description  %in% c("called_strike", "swinging_strike",
                                           "foul_tip", "bunt_foul_tip", "foul_bunt",
                                           "swinging_strike_blocked", "missed_bunt"),
                       1, 0))


# Batter Whiff
batter_stats_l <- lhp %>% 
  summarize(rate = mean(whiff),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_whiff = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bwhiff = (pitches / 300)*rate + ((300-pitches)/300)*pitch_whiff) %>% 
  mutate(pred_bwhiff = ifelse(pitches >= 300, rate, pred_bwhiff))


# Merging Whiff Prediction with LHP
lhp <- lhp %>% 
  left_join(select(batter_stats_l, batter, pitch_type, pred_bwhiff), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))

lhp <- lhp %>% 
  mutate(prev_pitch = ifelse(is.na(prev_pitch), "None", prev_pitch)) %>% 
  mutate(Count = paste0(balls, "-", strikes)) %>% 
  mutate(Count = as.factor(Count))

# Adding Proportional Distance
lhp <- lhp %>% 
  mutate(dist_x = plate_x/0.708333,
         center = (sz_top + sz_bot) / 2,
         dist_z = ifelse(plate_z >= center, 
                         (plate_z - center) / (sz_top - center), 
                         (plate_z - center) / (center - sz_bot)),
         dist_prop = sqrt(dist_z^2 + dist_x^2))


# Barrel
batter_stats2_l <- lhp %>% 
  summarize(rate = mean(Barrel),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_barrel = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bbarrel = (pitches / 300)*rate + ((300-pitches)/300)*pitch_barrel) %>% 
  mutate(pred_bbarrel = ifelse(pitches >= 300, rate, pred_bbarrel))

lhp <- lhp %>% 
  left_join(select(batter_stats2_l, batter, pitch_type, pred_bbarrel), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))


# Strike
batter_stats3_l <- lhp %>% 
  summarize(rate = mean(STRIKE),
                      pitches = n(),
            .by = c(batter, pitch_type)) %>% 
  mutate(pitch_strike = weighted.mean(rate, pitches),
         .by = pitch_type) %>% 
  mutate(pred_bstrike = (pitches / 300)*rate + ((300-pitches)/300)*pitch_strike) %>% 
  mutate(pred_bstrike = ifelse(pitches >= 300, rate, pred_bstrike))

lhp <- lhp %>% 
  left_join(select(batter_stats3_l, batter, pitch_type, pred_bstrike), 
             by = c("batter" = "batter", "pitch_type" = "pitch_type"))

```


```{r Pitcher Season Stats}
full <- rbind(rhp, lhp) %>% 
  mutate(Barrel = ifelse(Barrel == TRUE, 1, 0),
         whiff = ifelse(whiff == TRUE, 1, 0))

pitcher_list <- full %>% 
  group_by(player_id, hitter, pitch_type) %>% 
  summarize(n = n())

full <- full %>% 
  left_join(pitcher_list, by = c("player_id", "hitter", "pitch_type")) %>% 
  filter(n >= 100)

full_summary <- full %>% 
  group_by(player_id, player_name, pitch_type, p_throws, hitter) %>% 
  summarize("Pitches" = n(), 
            "Speed" = mean(pitch_speed, na.rm = TRUE), 
            "Spin Rate" = mean(release_spin_rate, na.rm = TRUE), 
            "X Movement" = mean(pfx_x, na.rm = TRUE), 
            "Z Movement" = mean(pfx_z, na.rm = TRUE),
            "Whiff Prop" = mean(whiff, na.rm = TRUE), 
            "Strike Prop" = mean(STRIKE, na.rm = TRUE),
            "Barrel Prop" = mean(Barrel, na.rm = TRUE)) %>% 
  mutate(Name = sapply(player_name, swap_names)) %>% 
  ungroup() %>% 
  select(Name, pitch_type:`Barrel Prop`)

# write.csv(full_summary, "Pitcher_Comps.csv")
```


### Four Seam Fastballs RHP to RHH

##### Whiff - Removed: pfx_x
##### Barrel - Removed: pfx_z, dist_prop, pfx_x, pfx_total, zone (multiple at 0.05-0.06)
##### Strike - Removed: pitch_speed, pfx_x, pfx_total

```{r FB RHP RHH, warning = FALSE, message = FALSE}
fb_r <- rhp %>% 
  filter(pitch_type == "FF",
         hitter == "R") %>% 
  mutate(Barrel = ifelse(Barrel == TRUE, 1, 0),
         whiff = ifelse(whiff == TRUE, 1, 0))


# Whiff ####

# Refined Model
fb_r_whiff <- glm(whiff ~ pitch_speed + pfx_z + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bwhiff,
              data = fb_r, family = binomial)

# Removed: pfx_x

# Model Evaluation
summary(fb_r_whiff)

fb_r_whiff_test <- augment(fb_r_whiff, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
   mutate(Response = "whiff",
         Hand = "right",
         pitch = "fastball")


rmse(fb_r_whiff_test$prediction, fb_r_whiff_test$whiff)



# Barrel ####

# Refined Model
fb_r_barrel <- glm(Barrel ~ pitch_speed + dist_x + dist_z + zone +
                  I(dist_x^2) + I(dist_z^2) + pred_bbarrel,
              data = fb_r, family = binomial)

# Removed: pfx_x, dist_prop, pfx_total, pfx_z, release_spin_rate

# Model Evaluation
summary(fb_r_barrel)

fb_r_barrel_test <- augment(fb_r_barrel, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "barrel",
         Hand = "right",
         pitch = "fastball")


rmse(fb_r_barrel_test$prediction, fb_r_barrel_test$Barrel)




# Strike ####

# Refined Model
fb_r_strike <- glm(STRIKE ~ pfx_z + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bwhiff,
              data = fb_r, family = binomial)

# Removed: pitch_speed, pfx_x, pfx_total

# Model Evaluation
# summary(fb_r_strike)

fb_r_strike_test <- augment(fb_r_strike, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>%
  mutate(Response = "strike",
         Hand = "right",
         pitch = "fastball")


rmse(fb_r_strike_test$prediction, fb_r_strike_test$STRIKE)

```



### Four Seam Fastballs RHP to LHH

##### Whiff - Removed: pfx_x, pfx_z
##### Barrel - Removed: release_spin_rate + pfx_total
##### Strike - Removed: pfx_x, pfx_z

```{r FB RHP LHH, warning = FALSE, message = FALSE}
fb_l <- rhp %>% 
  filter(pitch_type == "FF",
         hitter == "L") %>% 
  mutate(Barrel = ifelse(Barrel == TRUE, 1, 0),
         whiff = ifelse(whiff == TRUE, 1, 0))


# Whiff ####

# Refined Model
fb_l_whiff <- glm(whiff ~ pitch_speed + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bwhiff,
              data = fb_l, family = binomial)

# Removed: pfx_x, pfx_z

# Model Evaluation
# summary(fb_l_whiff)

fb_l_whiff_test <- augment(fb_l_whiff, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "whiff",
         Hand = "left",
         pitch = "fastball")


rmse(fb_l_whiff_test$prediction, fb_l_whiff_test$whiff)




# Barrel ####

# Refined Model
fb_l_barrel <- glm(Barrel ~ pitch_speed + pfx_x + zone + dist_x + 
                  dist_z + I(dist_x^2) + I(dist_z^2) +  pred_bbarrel,
              data = fb_l, family = binomial)

# Removed: release_spin_rate + pfx_total

# Model Evaluation
# summary(fb_l_barrel)

fb_l_barrel_test <- augment(fb_l_barrel, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "barrel",
         Hand = "left",
         pitch = "fastball")


rmse(fb_l_barrel_test$prediction, fb_l_barrel_test$Barrel)





# Strike ####

# Refined Model
fb_l_strike <- glm(STRIKE ~ pitch_speed + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bstrike,
              data = fb_l, family = binomial)

# Removed: pfx_x, pfx_z

# Model Evaluation
# summary(fb_l_strike)

fb_l_strike_test <- augment(fb_l_strike, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "strike",
         Hand = "left",
         pitch = "fastball")


rmse(fb_l_strike_test$prediction, fb_l_strike_test$STRIKE)
```



### Sliders RHP to RHH

##### Whiff - Removed: pitch_speed
##### Barrel - Removed: release_spin_rate, pfx_total, dist_prop, zone, pfx_x, pfx_z, pitch_speed
##### Strike - Removed: None

```{r SL RHP RHH, warning = FALSE, message = FALSE}
sl_r <- rhp %>% 
  filter(pitch_type == "SL",
         hitter == "R") %>% 
  mutate(Barrel = ifelse(Barrel == TRUE, 1, 0),
         whiff = ifelse(whiff == TRUE, 1, 0))


# Whiff ####

# Refined Model
sl_r_whiff <- glm(whiff ~ pfx_x + pfx_z + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bwhiff,
              data = sl_r, family = binomial)

# Removed: pitch_speed

# Model Evaluation
# summary(sl_r_whiff)

sl_r_whiff_test <- augment(sl_r_whiff, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "whiff",
         Hand = "right",
         pitch = "slider")


rmse(sl_r_whiff_test$prediction, sl_r_whiff_test$whiff)





# Barrel ####

# Refined Model
sl_r_barrel <- glm(Barrel ~ dist_x + dist_z +
                  I(dist_x^2) + I(dist_z^2) + pred_bbarrel,
              data = sl_r, family = binomial)

# Removed: release_spin_rate, pfx_total, dist_prop, zone, pfx_x, pfx_z, pitch_speed

# Model Evaluation
# summary(sl_r_barrel)

sl_r_barrel_test <- augment(sl_r_barrel, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "barrel",
         Hand = "right",
         pitch = "slider")


rmse(sl_r_barrel_test$prediction, sl_r_barrel_test$Barrel)





# Strike ####

# Refined Model
sl_r_strike <- glm(STRIKE ~ pitch_speed + pfx_x + pfx_z + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bstrike,
              data = sl_r, family = binomial)

# Removed: None

# Model Evaluation
# summary(sl_r_strike)

sl_r_strike_test <- augment(sl_r_strike, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "strike",
         Hand = "right",
         pitch = "slider")


rmse(sl_r_strike_test$prediction, sl_r_strike_test$STRIKE)


```




### Sliders RHP to LHH

##### Whiff - Removed: pfx_z, pfx_x, pfx_total
##### Barrel - Removed: pfx_total, dist_prop, release_spin_rate
##### Strike - Removed: pfx_z, release_spin_rate, pitch_speed

```{r SL RHP LHH, warning = FALSE, message = FALSE}
sl_l <- rhp %>% 
  filter(pitch_type == "SL",
         hitter == "L") %>% 
  mutate(Barrel = ifelse(Barrel == TRUE, 1, 0),
         whiff = ifelse(whiff == TRUE, 1, 0))


# Whiff ####

# Refined Model
sl_l_whiff <- glm(whiff ~ pitch_speed + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + release_spin_rate + pred_bwhiff,
              data = sl_l, family = binomial)

# Removed: pfx_z, pfx_x, pfx_total

# Model Evaluation
# summary(sl_l_whiff)

sl_l_whiff_test <- augment(sl_l_whiff, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "whiff",
         Hand = "left",
         pitch = "slider")

rmse(sl_l_whiff_test$prediction, sl_l_whiff_test$whiff)





# Barrel ####

# Refined Model
sl_l_barrel <- glm(Barrel ~ pitch_speed + pfx_x + pfx_z + zone + dist_x + dist_z +
                  I(dist_x^2) + I(dist_z^2) + pred_bbarrel,
              data = sl_l, family = binomial)

# Removed: pfx_total, dist_prop, release_spin_rate

# Model Evaluation
# summary(sl_l_barrel)

sl_l_barrel_test <- augment(sl_l_barrel, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "barrel",
         Hand = "left",
         pitch = "slider")


rmse(sl_l_barrel_test$prediction, sl_l_barrel_test$Barrel)







# Strike ####

# Refined Model
sl_l_strike <- glm(STRIKE ~ pfx_x + pfx_total + zone + dist_x + dist_z +
                  dist_prop + I(dist_x^2) + I(dist_z^2) + pred_bstrike,
              data = sl_l, family = binomial)

# Removed: pfx_z, release_spin_rate, pitch_speed

# Model Evaluation
# summary(sl_l_strike)

sl_l_strike_test <- augment(sl_l_strike, type.predict = "response") %>% 
  rename(prediction = `.fitted`) %>% 
  mutate(Response = "strike",
         Hand = "left",
         pitch = "slider")


rmse(sl_l_strike_test$prediction, sl_l_strike_test$STRIKE)
```


```{r Models, warning = FALSE, message = FALSE}
models <- rbind(
  tidy(fb_r_whiff) %>% mutate(Response = "whiff",
  Hand = "left", Pitch = "fastball"),
  tidy(fb_r_barrel) %>% mutate(Response = "barrel",
  Hand = "left", Pitch = "fastball"),
  tidy(fb_r_strike) %>% mutate(Response = "strike",
  Hand = "left", Pitch = "fastball"),
  
  tidy(fb_l_whiff) %>% mutate(Response = "whiff",
  Hand = "left", Pitch = "fastball"),
  tidy(fb_l_barrel) %>% mutate(Response = "barrel",
  Hand = "left", Pitch = "fastball"),
  tidy(fb_l_strike) %>% mutate(Response = "strike",
  Hand = "left", Pitch = "fastball"),
  
  tidy(sl_r_whiff) %>% mutate(Response = "whiff",
  Hand = "right", Pitch = "slider"),
  tidy(sl_r_barrel) %>% mutate(Response = "barrel",
  Hand = "right", Pitch = "slider"),
  tidy(sl_r_strike) %>% mutate(Response = "strike",
  Hand = "right", Pitch = "slider"),
  
  tidy(sl_l_whiff) %>% mutate(Response = "whiff",
  Hand = "left", Pitch = "slider"),
  tidy(sl_l_barrel) %>% mutate(Response = "barrel",
  Hand = "left", Pitch = "slider"),
  tidy(sl_l_strike) %>% mutate(Response = "strike",
  Hand = "left", Pitch = "slider"))


# write.csv(models, "models2.csv")
```



```{r Pred Averages, warning = FALSE, message = FALSE}
batter_preds_r <- batter_stats %>% 
  left_join(batter_stats2, by = c("batter", "pitch_type")) %>% 
  left_join(batter_stats3, by = c("batter", "pitch_type")) %>% 
  select(batter, pitch_type, pitches, pred_bwhiff, pred_bbarrel, pred_bstrike) %>% 
  mutate(hitter = "R")

batter_preds_l <- batter_stats_l %>% 
  left_join(batter_stats2_l, by = c("batter", "pitch_type")) %>% 
  left_join(batter_stats3_l, by = c("batter", "pitch_type")) %>% 
  select(batter, pitch_type, pitches, pred_bwhiff, pred_bbarrel, pred_bstrike) %>% 
  mutate(hitter = "L")

pred_means <- rbind(batter_preds_r, batter_preds_l) %>% 
  group_by(hitter, pitch_type) %>% 
  summarize(whiff_mean = weighted.mean(pred_bwhiff, pitches, na.rm = TRUE),
            barrel_mean = weighted.mean(pred_bbarrel, pitches, na.rm = TRUE),
            strike_mean = weighted.mean(pred_bstrike, pitches, na.rm = TRUE))

# write.csv(pred_means, "pred means.csv")
```


